### 2016-05-09 presentation/main.tex
separated proof of exploring pebble machine into lemmata

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}

@@

\title{A Grimm idea}
\subtitle{Exploring graphs with pebbles}
\author{Christoph Welzel}
\institute{Logik und Theorie diskreter Systeme, RWTH Aachen}

\begin{document}
\maketitle
\begin{frame}
  \frametitle{Motivation}
  \begin{itemize}
    \item Traversing Brobdingnagian graphs
    \item[$\rightarrow$] Web crawlers
    \item Agent moves over vertices along edges
    \item[$\Rightarrow$] Memory efficient agents
  \end{itemize}
\end{frame}

\section{Graph traversial}
\begin{frame}
  \frametitle{Agents}
  \begin{quotation}
    Und als der volle Mond aufgestiegen war, so nahm Hänsel sein
    Schwesterchen an der Hand und ging den Kieselsteinen nach, die schimmerten
    wie neu geschlagene Batzen und zeigten ihnen den Weg.
  \end{quotation}
  \vspace{-0.5cm}
  \flushright{Brüder Grimm}
  \vspace{-0.5cm}
  \flushright{\emph{Hänsel und Gretel}}
  \begin{itemize}
    \item Agent carries set of markers $\rightarrow$ pebbles
    \item Leave pebbles on vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Brobdignagian graphs}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Indistinguishable vertices
      \item Bounded degree ($\Delta$)
      \item Edges can locally enumerated $\left\{0,\dots,\Delta-1\right\}$
        (Port)
      \item Agent is aware of:
        \begin{enumerate}
          \item Degree  of vertex
          \item Pebbles placed on vertex
          \item Pebbles carried by agent
          \item Port    agent entered vertex
        \end{enumerate}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Describe movement of agent by relative turns
      \item $e_{1},\dots, e_{n}$
      \item For a vertex $v$ with degree $d_{v}$ which is entered on port $p$
        is left through port $p + e_{i}\mod p$
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{theorem}
    A exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ produces a closed walk.
  \end{theorem}
  \begin{itemize}
    \item Introduce triple $(v,l,i)$ with vertex $v$ is entered
      through port $l$ and $e_{i}$ is the next turn
    \item Note that there is exactly one $(v,l,i) \leadsto (v',l',i')$
    \item Furthermore assume $t_{1} = (v',l',i') \leadsto (v,l,i)$ and
      $t_{2} = (v'',l'',i'') \leadsto (v,l,i)$:
      \begin{itemize}
        \item $t_{1}$ and $t_{2}$ both entered $v$ through $l$: $v' = v''$
        \item $t_{1}$ and $t_{2}$ lead that $i$ is the next turn: $i' = i''$
        \item $l' + e_{i'} \mod d_{v'} = l'' + e_{i''} \mod d_{v''}$ and
          $l',l''\leq d_{v'} = d_{v''}$: $l' = l''$
      \end{itemize}
    \item Only finite triples, thus projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \alt<+>{
    \begin{theorem}[Reingold]
      There is an algorithm $A$ with $\mathcal{O}(n)$ space requirement
      producing a universal exploration sequence for every regular graph with
      $n$ vertices.
    \end{theorem}
  }{
    \begin{theorem}
      Thus, there is $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$ that moves
      along a closed walk with $n$ distinct vertices (or explores the whole
      graph).
    \end{theorem}
  }
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> Establish 3-regularity
        \begin{itemize}
          \item Transform $v$ to $3\cdot d_{v}$ distinct vertices
          \item TODO: ADD BOUNDARY
        \end{itemize}
      \item<5-> Use $A$ and take first $k = 3n\cdot c$ ($c$ number of
        configurations $M_{A}$)
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<4->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}

@@

\title{Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles}
\subtitle{A Grimm idea}
\author{Christoph Welzel}
\institute{Logik und Theorie diskreter Systeme, RWTH Aachen}

\begin{document}
% \tikzexternalize
\maketitle
\begin{frame}
  \frametitle{Motivation}
  \begin{itemize}
    \item Exploration of huge finite graphs with bounded degree by agents
    \item Agents are entities which are located on a vertex and can move along
      an edge
    \item Agents can drop markers on vertices
    \item Agent \emph{explores} a graph by deterministically visiting all of
      its vertices
  \end{itemize}
\end{frame}

\section{Preliminaries}
\begin{frame}
  \frametitle{Huge graphs}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Finite, but huge
      \item Undirected
      \item Indistinguishable vertices
      \item Bounded degree ($\Delta$)
      \item Edges can locally be enumerated as $0,\dots,\Delta-1$
        (Port)
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble-machines
        \begin{itemize}
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item $P$: set of pebbles
          \item $m$: length of the used memory tape
          \item $\din$: formalisation of agent entering
          \item $\dout$ formalisation of agent leaving
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item As an agent moves on a vertex it observes:
        \begin{itemize}
          \item Pebbles it carries
          \item Pebbles lying on the vertex
          \item Degree of vertex
        \end{itemize}
      \item Agent executes a computation on a vertex to determine:
        \begin{itemize}
          \item Port to leave
          \item Pebbles to leave behind
          \item Pebbles to carry with it
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Describe movement of agent by relative movements:
        \begin{itemize}
          \item An exploration sequence has the form $e_{1},\dots, e_{n}$
          \item For a vertex $v$ with degree $d_{v}$ which is entered on port
            $p$ is left through port $p + e_{i}\mod d_{v}$
          \item An exploration sequence is called \emph{universial} for a class
            of graphs if following it explores any graph of that class
          \item<2->[$\rightarrow$] Example:
            $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
            \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
            \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
            \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[decorate,decoration={snake,amplitude=0.3cm, segment length=1cm},->]
        (d1) to (d2);
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}
\begin{frame}
  \frametitle{Closed walk}
  \begin{lemma}
    An agent following an exploration sequence of the form
    $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
  \end{lemma}
  \begin{itemize}
    \item Examine triples $(v,\ell,i)$:
      \begin{itemize}
        \item $v$: current vertex
        \item $\ell$: port agent entered $v$ through
        \item $i$: next step of the exploration sequence
      \end{itemize}
    \item Deterministically $(v,\ell,i) \leadsto (v',\ell',i')$
    \item If $t_{1} = (v',\ell',i') \leadsto (v,\ell,i)$ and
      $t_{2} = (v'',\ell'',i'') \leadsto (v,\ell,i)$: $t_{1} = t_{2}$
    \item Only finite triples: projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Universal exploration of $3$-regular graphs}
  \begin{lemma}
    There exists an $\mathcal{O}(\log n)$-space algorithm producing a universal
    exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any $3$-regular graph
    with at most $n$ vertices.
  \end{lemma}
  \begin{itemize}
    \item Let $\mathcal{M}$ be the Turing machine from Reingold's Theorem
    \item The exploration sequence computed by $\mathcal{M}$ is universial for
      3-regular graphs with at most $n$ vertices
      \alt<1-1>{
        \begin{itemize}
          \item Assume $3$-regular $G_{0}$ with $n_{0} < n$ vertices is not
            explored
          \item Then there is a vertex that is not traversed
          \item Repeatedly apply \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item Until there are $n$ vertices which contradicts Reingold's
            Theorem
        \end{itemize}
      }{
      \item Produce a periodic universal exploration sequence
        $(e_{1},\dots,e_{k})^{\ast}$
        \begin{itemize}
          \item Construct $\mathcal{M}'$ which simulates $\mathcal{M}$ for
            $3n\cdot c$ steps where $c$ is the number of configurations of $M$
          \item Output $e = (e_{1},\dots,e_{k})^{\ast}$
          \item Assume $e$ is not universal; then it loops in $V'$ with
            $|V'|<n$
          \item $\mathcal{M}$ has to enter a vertex twice with the same
            configuration
          \item Contradicts universality of the exploration sequence of
            $\mathcal{M}$
        \end{itemize}
      }
  \end{itemize}
  \pause
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<2-> Establish 3-regularity
        \begin{itemize}
          \item Transform $v$ to $3\cdot d_{v}$ distinct vertices
        \end{itemize}
      \item<4-> Use Reingold's algorithm and take first $k = 3n\cdot c$ ($c$ number of
        configurations $M_{A}$)
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<3->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

### 2016-05-09 tikz/settings.tex
separated proof of exploring pebble machine into lemmata

\usetikzlibrary{patterns, arrows, cd, positioning, backgrounds, fit, decorations.pathmorphing, shapes.misc, calc, matrix, decorations.pathreplacing}
\tikzset{%
  node/.style={draw, minimum size=12mm, inner sep=0mm, circle, fill=gray!50},
  edge/.style={thick},
  move/.style={->},
  brace/.style={
    decorate,
    decoration={brace,amplitude=10pt,mirror},
  },
  table nodes/.style={
    rectangle,
    draw=black,
    align=center,
    minimum height=7mm,
    text depth=0.5ex,
    text height=2ex,
    inner xsep=0pt,
    outer sep=0pt,
    inner sep=0pt
  },
  table/.style={
    matrix of nodes,
    row sep=-0.5pt,
    column sep=-0.5pt,
    nodes={
      table nodes
    }
  }
}

>>>>

\usetikzlibrary{patterns, arrows, cd, positioning, backgrounds, fit, decorations.pathmorphing, shapes.misc, calc, matrix, decorations.pathreplacing, external}
\tikzset{%
  node/.style={draw, minimum size=12mm, inner sep=0mm, circle, fill=gray!50},
  edge/.style={thick},
  move/.style={->},
  brace/.style={
    decorate,
    decoration={brace,amplitude=10pt,mirror},
  },
  table nodes/.style={
    rectangle,
    draw=black,
    align=center,
    minimum height=7mm,
    text depth=0.5ex,
    text height=2ex,
    inner xsep=0pt,
    outer sep=0pt,
    inner sep=0pt
  },
  table/.style={
    matrix of nodes,
    row sep=-0.5pt,
    column sep=-0.5pt,
    nodes={
      table nodes
    }
  }
}
